---
title: "r_data_types"
format: html
---
## Check data types
What is the class of the vector? class()
What type of variable does it store? typeof()
Access the 3rd element and store as vec_1_e3
Access the 1st element and store as vec_1_e1
Access the 5th through 7th elements and store as vec_1_e5to7
Reassign vec_1 as a character using as.character, stored as vec_1_char. What does the output look like?
```{r}
vec_1 <- c(2, 5, 9, 10, 8, 12, 1, 0) #creating vector
class(vec_1) #numeric
typeof(vec_1) #double

vec_1_e3 <- vec_1[3]
vec_1_e3

vec_1_e1 <- vec_1[1]
vec_1_e1

vec_1_e5to7 <- vec_1[5:7]
vec_1_e5to7

vec_1_char <- as.character(vec_1) # Reassign to character vector
vec_1_char #everything is in quotation marks

vec_1_logical <- as.logical(vec_1) # Reassign to logical vector
vec_1_logical #True False statement- all true except last one
vec_1
```


## Create a vector called vec_2
vec_2 should contained named elements, where town = "Santa Barbara, location = "Rincon", swell = "south"

Take a look at what you’ve made
What is the class of vector elements? class()
What is the length of vec_2?
Access the 2nd element by name and store as vec_2_e2
```{r}
vec_2 <- c(town = "Santa Barbara", #names elements
           location = "Rincon",
           swell = "south")
vec_2
class(vec_2) #character
length(vec_2) # 3

vec_2_e2 <- vec_2[2]
vec_2_e2
```

## Create a data frame in R
Write code to create a data frame called df_1 that looks like this:

  region     species count
1      A       otter    12
2      B great white     2
3      A    sea lion    36
4      D  gray whale     6
Return the class of the entire data frame
Return the class of the species column
Find the maximum value of the count() column, store as max_count
```{r}
df_1 <- data.frame(region = c("A", "B", "C", "D"),
                   species = c("otters", "great white", "sea lion", "gray whale"),
                   count = c(12, 2, 36, 6))
class(df_1) #data.frame
class(df_1$species) #character
max_count <- max(df_1$count) #36
```

